---
title: "Segmentation & Profiling Project"
author: "Daniel Jackson"
date: "`r Sys.Date()`"
output: word_document
---

# Table of Contents
Introduction. 
Summary of Findings.  
Detailed Technical Section 

# Introduction
This segmentation and profiling section of the project will be a continuation of our data due diligence project. As we did in the data due diligence project, we will acting as a marketing analytics manager for a telecommunications company. After cleaning the customer data set and creating a subset of the original data, we are now tasked with developing customer segmentation that can support effective, and economically sound, customer retention efforts.
To accomplish this goal, we will be using two of the three segmentation techniques: rules-based, supervised, or unsupervised (We need to figure out which two that we will be doing and justify why we are using each method).
In this project, we will provide a summary of our key findings and recommendations that we will make as marketing analytics manager. We will highlight notable insights and provide proper conclusions and recommendations based on our findings. 
Then, we will run through a technical section describing the segmentation methods that we chose, highlight detailed findings, and use visualizations created in Tableau to depict the individual customer segments that we created. 
We will begin by reading in our data subset that we created in our previous data due diligence project. 

``` {r libraries, echo = FALSE, results = 'hide', message = FALSE}
library(dplyr)
library(tidyr)
```

```{r setup, include = TRUE, eval = FALSE}
# Packages used
library(dplyr)
library(tidyr)

# Read in data
sub_df = read.csv("combined_customer_data.csv")

# Check dimension of data frame
dim(sub_df)
# 4347 observations
# 11 columns

# Let us remove X column
sub_df = sub_df[, -which(names(sub_df) == "X")]

dim(sub_df)
# 4347 observations
# 10 columns

str(sub_df)
# Let us change numberpets, household size to numeric vectors and change homeowner to Yes or No (1 for yes, 0 for no)
sub_df = sub_df %>%
  mutate(
    numberpets = as.numeric(numberpets),
    householdsize = as.numeric(householdsize),
    homeowner = ifelse(homeowner == 1, "Yes", "No")
  )

str(sub_df)
```

```{r Rules-Based Segmentation 1, include = TRUE, eval = FALSE}
# Check min, max and average
min(sub_df$log_tot_value)
# 2.03 min value
min_val = min(sub_df$log_tot_value)

max(sub_df$log_tot_value)
# 9.95 max value
max_val = max(sub_df$log_tot_value)

mean(sub_df$log_tot_value)
# 6.46 average value
mean_val = mean(sub_df$log_tot_value)

# 1/3 threshold
one_third_val = ((1/3) * max_val)

# 2/3 threshold
two_third_val = (2 * one_third_val)


# Let us create customer_value qualitative variable using our breaks that we created using the cut function
sub_df$customer_value = cut(sub_df$log_tot_value, breaks=c(0, one_third_val, two_third_val, max_val), labels=c("Low", "Medium", "High"))
unique(sub_df$customer_value)
# Check if any NA values in customer_value variable
any(is.na(sub_df$customer_value))
# FALSE

table(sub_df$customer_value)

# Now let us combine this rule with a few variables
```

```{r Rules-Based Segmentation 2, include = TRUE, eval = FALSE}
# Region
unique(sub_df$region)
sub_df = sub_df %>% mutate(reg_segment = case_when(
      customer_value == "Low" & region == "northeast" ~ 1,
      customer_value == "Medium" & region == "northeast" ~ 2,
      customer_value == "High" & region == "northeast" ~ 3,
      customer_value == "Low" & region == "southeast" ~ 4,
      customer_value == "Medium" & region == "southeast" ~ 5,
      customer_value == "High" & region == "southeast" ~ 6,
      customer_value == "Low" & region == "west" ~ 7,
      customer_value == "Medium" & region == "west" ~ 8,
      customer_value == "High" & region == "west" ~ 9,
      customer_value == "Low" & region == "southwest" ~ 10,
      customer_value == "Medium" & region == "southwest" ~ 11,
      customer_value == "High" & region == "southwest" ~ 12,
      customer_value == "Low" & region == "midwest" ~ 13,
      customer_value == "Medium" & region == "midwest" ~ 14,
      customer_value == "High" & region == "midwest" ~ 15
    ))

# 1 = low and northeast
# 2 = medium and northeast
# 3 = high and northeast
# 4 = low and southeast
# 5 = medium and southeast
# 6 = high and southeast
# 7 = low and west
# 8 = medium and west
# 9 = high and west
# 10 = low and southwest
# 11 = medium and southwest
# 12 = high and southwest
# 13 = low  and midwest
# 14 = medium and midwest
# 15 = high and midwest

# Check for any NA values in segment variable
any(is.na(sub_df$reg_segment))
# FALSE

# Summarize results of each segment
sub_df %>% 
  add_count(reg_segment) %>%
  group_by(reg_segment,n) %>% 
  summarise_all("mean")
```

```{r Rules-Based Segmentation 3, include = TRUE, eval = FALSE}
# Remove qualitative variables
sub_df %>% 
  select(-c("region","homeowner","jobcategory", "carbrand", "customer_value")) %>% 
  add_count(reg_segment) %>%
  group_by(reg_segment,n) %>% 
  summarise_all("mean")
```

```{r Rules-Based Segmentation 4, include = TRUE, eval = FALSE} 
# Homeowner
unique(sub_df$homeowner)
sub_df = sub_df %>% mutate(home_segment = case_when(
      customer_value == "Low" & homeowner == "Yes" ~ 1,
      customer_value == "Medium" & homeowner == "Yes" ~ 2,
      customer_value == "High" & homeowner == "Yes" ~ 3,
      customer_value == "Low" & homeowner == "No" ~ 4,
      customer_value == "Medium" & homeowner == "No" ~ 5,
      customer_value == "High" & homeowner == "No" ~ 6
    ))

# 1 = low and homeowner
# 2 = medium and homeowner
# 3 = high and homeowner
# 4 = low and non-homeowner
# 5 = medium and non-homeowner
# 6 = high and non-homeowner

# Summarize results of each segment
sub_df %>% 
  add_count(home_segment) %>%
  group_by(home_segment,n) %>% 
  summarise_all("mean")
```

```{r Rules-Based Segmentation 5, include = TRUE, eval = FALSE}
# Remove qualitative variables
sub_df %>% 
  select(-c("region","homeowner","jobcategory", "carbrand", "customer_value")) %>% 
  add_count(home_segment) %>%
  group_by(home_segment,n) %>% 
  summarise_all("mean")
```

```{r Rules-Based Segmentation 6, include = TRUE, eval = FALSE} 
# Job Category
unique(sub_df$jobcategory)
sub_df = sub_df %>% mutate(job_segment = case_when(
      customer_value == "Low" & jobcategory == "Professional" ~ 1,
      customer_value == "Medium" & jobcategory == "Professional" ~ 2,
      customer_value == "High" & jobcategory == "Professional" ~ 3,
      customer_value == "Low" & jobcategory == "Sales" ~ 4,
      customer_value == "Medium" & jobcategory == "Sales" ~ 5,
      customer_value == "High" & jobcategory == "Sales" ~ 6,
      customer_value == "Low" & jobcategory == "Labor" ~ 7,
      customer_value == "Medium" & jobcategory == "Labor" ~ 8,
      customer_value == "High" & jobcategory == "Labor" ~ 9,
      customer_value == "Low" & jobcategory == "Agriculture" ~ 10,
      customer_value == "Medium" & jobcategory == "Agriculture" ~ 11,
      customer_value == "High" & jobcategory == "Agriculture" ~ 12,
      customer_value == "Low" & jobcategory == "Service" ~ 13,
      customer_value == "Medium" & jobcategory == "Service" ~ 14,
      customer_value == "High" & jobcategory == "Service" ~ 15,
      customer_value == "Low" & jobcategory == "Crafts" ~ 16,
      customer_value == "Medium" & jobcategory == "Crafts" ~ 17,
      customer_value == "High" & jobcategory == "Crafts" ~ 18
    ))

# 1 = low and professional
# 2 = medium and professional
# 3 = high and professional
# 4 = low and sales
# 5 = medium and sales
# 6 = high and sales
# 7 = low and labor
# 8 = medium and labor
# 9 = high and labor
# 10 = low and agriculture
# 11 = medium and agriculture
# 12 = high and agriculture
# 13 = low  and service
# 14 = medium and service
# 15 = high and service
# 16 = low and crafts
# 17 = medium and crafts
# 18 = high and crafts

# Summarize results of each segment
sub_df %>% 
  add_count(job_segment) %>%
  group_by(job_segment,n) %>% 
  summarise_all("mean")
```

```{r Rules-Based Segmentation 7, include = TRUE, eval = FALSE}
# Remove qualitative variables
sub_df %>% 
  select(-c("region","homeowner","jobcategory", "carbrand", "customer_value")) %>% 
  add_count(job_segment) %>%
  group_by(job_segment,n) %>% 
  summarise_all("mean")
```

```{r Rules-Based Segmentation 8, include = TRUE, eval = FALSE} 
# Car Brand
unique(sub_df$carbrand)
sub_df = sub_df %>% mutate(car_segment = case_when(
      customer_value == "Low" & carbrand == "Domestic" ~ 1,
      customer_value == "Medium" & carbrand == "Domestic" ~ 2,
      customer_value == "High" & carbrand == "Domestic" ~ 3,
      customer_value == "Low" & carbrand == "Foreign" ~ 4,
      customer_value == "Medium" & carbrand == "Foreign" ~ 5,
      customer_value == "High" & carbrand == "Foreign" ~ 6
    ))

# 1 = low and domestic
# 2 = medium and domestic
# 3 = high and domestic
# 4 = low and foreign
# 5 = medium and foreign
# 6 = medium and foreign

# Summarize results of each segment
sub_df %>% 
  add_count(car_segment) %>%
  group_by(car_segment,n) %>% 
  summarise_all("mean")
```

```{r Rules-Based Segmentation 9, include = TRUE, eval = FALSE}
# Remove qualitative variables
sub_df %>% 
  select(-c("region","homeowner","jobcategory", "carbrand", "customer_value")) %>% 
  add_count(car_segment) %>%
  group_by(car_segment,n) %>% 
  summarise_all("mean")
```

Now that we have created 4 different rule-based segments, let us use Tableu to visualize each to help us select best rules-bases segment to represent our customers.

```{r export sub_df to csv file, include = TRUE, eval = FALSE}
sub_df %>% 
  write.csv("segmentation_df.csv")
```

```{r Unsupervised 1, include = TRUE, eval = FALSE}
# Let us re-read data in
sub_df = read.csv("combined_customer_data.csv")

sub_df = sub_df[, -which(names(sub_df) == "X")]

sub_df = sub_df %>%
  mutate(
    numberpets = as.numeric(numberpets),
    householdsize = as.numeric(householdsize),
    homeowner = ifelse(homeowner == 1, "Yes", "No")
  )

# Now, let us create subset of data that only has numeric variables
num_df = sub_df[sapply(sub_df, is.numeric)]

# Let us scale each variable before we perform K-means clustering
# Assuming num_df contains only numeric variables
scaled_df = as.data.frame(scale(num_df))
```

```{r Unsupervised 2, include = TRUE, eval = FALSE}
# Now, let us take a look and see what K should equal for our clustering segmention
ks = 1:12
tot_within_ss = sapply(ks, function(k) {
    set.seed(1223)
    cl = kmeans(scaled_df, k)
    cl$tot.withinss
})
plot(ks, tot_within_ss, type = "b")
```

```{r Unsupervised 3, include = TRUE, eval = FALSE}
# Based on plot, we see biggest drop off from K = 1 to K = 2.
# We see that once K = 6, the drop in tot_within_ss value tends to drop relatively constantly as the value of K increases. Therefore, let us use K = 6 for our segmentation.

set.seed(1223)
num_clusters = 6
kclust = kmeans(scaled_df, centers = num_clusters, nstart = 10)

#add segments to original dataset
sub_df$kmeans_six = as.factor(kclust$cluster)
```

```{r Unsupervised 4, include = TRUE, eval = FALSE}
sub_df %>% 
  select(-c("region","homeowner", "jobcategory", "carbrand")) %>% 
  add_count(kmeans_six) %>%
  group_by(kmeans_six,n) %>% 
  summarise_all("mean")
```

```{r Unsupervised 5, include = TRUE, eval = FALSE}
# Add K = 5 and K = 7 segments to data set so we can compare them visually
# K = 5
set.seed(1223)
num_clusters = 5
kclust = kmeans(scaled_df, centers = num_clusters, nstart = 10)

#add segments to original dataset
sub_df$kmeans_five = as.factor(kclust$cluster)

sub_df %>% 
  select(-c("region","homeowner", "jobcategory", "carbrand")) %>% 
  add_count(kmeans_five) %>%
  group_by(kmeans_five,n) %>% 
  summarise_all("mean")

# K = 7
set.seed(1223)
num_clusters = 7
kclust = kmeans(scaled_df, centers = num_clusters, nstart = 10)

#add segments to original dataset
sub_df$kmeans_seven = as.factor(kclust$cluster)

sub_df %>% 
  select(-c("region","homeowner", "jobcategory", "carbrand")) %>% 
  add_count(kmeans_seven) %>%
  group_by(kmeans_seven,n) %>% 
  summarise_all("mean")
```

```{r Unsupervised 6, include = TRUE, eval = FALSE}
# Export sub_df to CSV file again for unsupervised segmentation
sub_df %>% 
  write.csv("unsup_segmentation_df.csv")
```